Effects of rotation and magnetic braking on Li abundances for solar-type stars

### 2019-08-21 project/overleaf/paper.tex
Paper v3.2

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the hydrostatic effects of rotation on Li abundances are studied by simulating several grids of PMS \& MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB one (with magnetic field intensities ranging between 3.5 and 5.5 G) and the data obtained from the two set of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in star evolutionary models, impacts also the effective temperature ($T_{eff}$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the field intensity: the higher it is, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. Another important result of this study is that MB effect must be taken into consideration during PMS must if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

\section{Introduction}
Despite decades of theoretical efforts, a coherent explanation cannot be found from the models for the discrepancies found in the comparisons between the abundance of Li for stars belonging to clusters of different ages and that are also in the same evolutionary state, either in the PMS or in the MS. Additionally, these theoretical models are not able to explain the abundances detected in the late stages of the MS \citep{Tschape2001}.\par

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $T_{Li} \approx 2.5 x 10^6\; K$, and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. The Sun in particular and the solar-type stars in general are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $T_{Li}$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS) when the convection zone retreats and the temperature at BCZ is cooler than $T_{Li}$. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $T_{Li}$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances, in addition to showing a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this same period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the convective envelope, from its lower limit to the star surface. However, different abundances of Li have been observed for different star populations (\citet{Somers2014} and references therein).\par

From approximately the 1920s of the last century it has been accepted that rotation is capable of inducing the mixing of the elements present inside the stars. The studies of Eddington (1925) and Vogt (1925) had revealed the fact that the southern circulation must necessarily arise in those stars in rotation as a consequence of the von Zeipel paradox: a star cannot be simultaneously in thermal and hydrostatic equilibrium if its rotation speed depends exclusively on its radius. The solution to this depends on the existence of southern circulations. At the same time, this type of instability implies an impact on the convection mechanisms not considered by the standard models. Today it is known that in combination with this type of instability exists a catalogue of others that appear as a consequence of assuming models in rotation \citep{Maeder2003a}, especially the differential rotation as a function of the radius of the star or the deviation of the spherical symmetry. Although initially these effects could be candidates to explain the observations, the fact is that the Sun is a slow rotator, meaning slow rotation that one for which the maximum centrifugal force is much smaller than the force of gravity. Furthermore, its deviation of the measured spherical symmetry is of the order of $10^{-5}$ and the time scale for the southern circulation is of the order of $10^{12}$ years, a time period well above the nuclear combustion time scale \citep{Pinsonneault1997}. In view of these findings, it seems evident that there are still open questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models.\par

The impact of rotation on PMS and on Li depletion for solar-type stars has already extensively debated in the past (\citet{Pinsonneault1997}, \citet{Jeffries2004}, \citet{Somers2014}). However, these previous studies focused mainly on the hydrostatic effects and didn't consider the influence of a coupling between the star magnetic field and its possible spin-down effect. It's thought that AML has a direct influence in the mixing processes and that transport of the momentum could be produced by a number of relevant mechanisms: mass loss, magnetic fields and waves (g-modes). Concerning the two latter, waves and magnetic fields have the property that they generally transmit angular momentum (AM) much more effectively than they induce mixing \citep{Denissenkov2007}. As result, the efficiency of these processes is most easily tested by their impact on AM evolution. Of particular interest for this work is the impact caused by the magnetic field of the star in the AML, more specifically the effect produced on it by MB as a consequence of the solidary rotation that produces on the particles that constitute the solar wind and the torque derived from this coupling. These magnetic fields would originate in the limits between the radiative and convective zones of the star, in the so-called tachoclines, due to the rotational difference between both zones \citep{Guerrero2016}. A proper consideration of the interaction between rotation and magnetic fields is therefore essential. We have adapted the one-dimensional stellar evolution code used in this work to model MB and included its effect on the equations for the evolution of the AM distribution and stellar structure. This produces a more complete, though still reasonably simple, model for AM evolution. It's worth to mention that in this study we don't only focused in investigating the effects of rotation on the Li content during PMS and MS but also on the indirect role played by the MB on the Li destruction as a result of its influence on the rotational history of solar-type stars.\par

\subsection{Effects of rotation}
Rotation may affect the equations of stellar structure in four different ways according to \citet{Endal1976}:
%\renewcommand{\theenumi}{\arabic{enumi}.}
\begin{enumerate}
    \item The centrifugal force causes the local effective gravity ($g_{eff}$) associated with a point of the star not located on the axis of rotation to be reduced.
    \item In general, the centrifugal force will not be parallel to the gravitational force and, as a consequence of this, the equipotential layers will cease to be spherical.
    \item As the von Zeipel theorem states, the radiative flux in a uniformly rotating star is proportional to the $g_{eff}$, therefore the radiative flux is not constant on an equipotential surface.
    \item Rotation can suppress certain mechanisms that trigger convection, indirectly affecting mixing processes.
\end{enumerate}

Observational data support the fact that stars are born with different angular velocities even if they come from the same interstellar cloud and that throughout evolution they undergo a deceleration process \citep{Skumanich1972} in a time period between 0.5 and 5 Gyr \citep{Somers2014a}. This range of different velocities implies that the mixing mechanisms induced by rotation vary their efficiency and effectiveness and therefore have a direct influence on the abundances of the different chemical elements observed on the surface of stars.  This fact, together with the differential degree of rotation to which the limits between the radiative and convective zones of the star are subjected, is of particular interest for our study since Li, as mentioned in the introduction, is destroyed at a low temperature. Therefore, variations in the intensity of the mixing processes can cause convective zones to extend to interior zones with temperatures higher than $T_{Li}$. The deceleration process to which a star is subjected throughout its evolution and the direct link that the rotation has with the mixing processes also offer a natural and consistent explanation for the fact that the destruction of Li is more accentuated or efficient in young stars than in old stars \citep{Sestito2005}. In a similar way, stellar rotation rates show a large dispersion at ZAMS \citep{Stauffer1984}, providing in this way a consistent argument for explaining the variance of Li abundances among stars which have a similar mass and were formed within the same cloud. Although this approach is not a novelty at all, it is no less true that in those studies based on it the effect caused by MB over the rotation on the PMS and MS is often overlooked. \par

@@

In the remainder of this paper, we describe a physically sound yet computationally simple model extension to Modules for Experiments in Stellar Astrophysics star evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar wind. In its simplest form, the implementation of the model requires the specification of both the surface magnetic field strength ($B$) and the initial  rotation rate ($\Omega$) at star surface. Our simulations trace the rotational history and A(Li) of a $1\, M_{\sun}$ star for a variety of initial values for $B$ and $\Omega$.\par

We define here also some conventions and assumptions adopted in the paper. For the models presented in this paper we adopt solar-scaled abundances and assumes the \citet{Asplund2009} protosolar birth cloud metallicity ($Z = Z_{\odot, proto} = 0.0142$) not the current photospheric metallicity ($Z = Z_{\odot, photosphere} = 0.0134$) as the reference value. We also adopt the following nominal values to express stellar properties values in SI units $R_{\odot} = 6.957x10^{10}\, cm$ and $M_{\odot} = 1.988x10^{33}\, g$ which are consistent with IAU resolutions \citep{Mamajek2015}. For a detailed description of the physics adopted in this paper, refer to \citet{Choi2016}. That work, in particular its section 4.1, has been used as a starting point for ours regarding the parameterization of the MESA project which is calibrated to reproduce the measured values on the solar surface.\par

@@

The mass of stars plays a determining role in the existence or not of an external convective envelope. In turn, the presence of this convective region conditions the evolution of magnetic fields. Massive stars do not have external convective zones although they can expose a convective core. In these cases, the dominant type of energy transport in the outer zones of the star is radiative. On the other hand, the less massive stars, such as the Sun, have convective covers and a radiative core. In both types of stars strong magnetic fields can be generated in the tachocline but with notable differences. In the case of massive stars, these magnetic fields can survive for a long time but not reach the surface of the star. On the other hand in less massive stars these magnetic fields, although they are able to reach the surface of the star, decay quickly in time, in a few hundred years or even only in a few decades \citep{Chabrier2006}, so it is necessary a mechanism for generating magnetic field, e.g. a dynamo type, continued in time to explain its existence. The way in which this dynamo process is able to generate magnetic fields depends on different parameters and to this day this mechanism is still not fully accepted and is subject of controversy, e.g. \citep{Charbonneau2010}.\par

\subsubsection{Internal Magnetic Fields}
Our Sun is an active star as far as the magnetic field is concerned. According to the theoretical models this magnetic field is generated by a dynamo type effect in the tachocline, the intermediate zone between the radiative core and the CZ \citep{Aschwanden2014}, and has an intensity of approximately $B\approx10^5\, G$. Additionally, these models predict that because of the differential rotation existing between the core and CZ, the lines of the magnetic field that at the moment of their generation have a poloidal orientation will be affected by the differential rotation and forced to adopt a toroidal typology. This change of typology in the magnetic field, from poloidal to toroidal, is known as the $\omega$-effect. On the other hand, we have that the opposite case can also occur, that is to say, starting from a toroidal magnetic field to obtain a poloidal configuration. This would be possible, on the one hand, thanks to the convective movements of material that are produced in the CZ of the star and, on the other, to assume that this material behaves as a perfect conductor. This second condition allows to assume that the magnetic field lines are "frozen" in the conductive fluid and therefore must move in solidarity with it. Under these circumstances the magnetic field lines will be pushed to higher areas of the CZ as the material moves vertically and, due to the combined effect of stellar rotation plus Coriolis forces, the magnetic lines will be twisted and give rise to a poloidal magnetic field. This is what is known as the $\alpha$-effect.\par

A number of theoretical models have been proposed based on $\alpha$ and $\omega$ effects as an explanation to the generation of magnetic fields, among which is the so-called $\alpha\omega$ or also known as dynamo effect which was proposed by \citet{Spruit2002}. By means of this model it is possible to explain the change in polarity of the magnetic field that occurs in the Sun every 11 years, completing a full cycle every 22 years. It's also worth to mention that a theoretical debate on this topic is still ongoing \citep{Denissenkov2007} and with regards to MESA, an implementation according to the work of Spruit-Tayler has been implemented \citep{Paxton2013}.\par

\subsubsection{Surface Magnetic Fields} \label{surf_mf}
The intensity of the magnetic field measured on the Sun's surface is about $B\approx1\, G$, twice the value of the Earth's one. Additionally its topology is varied and complex. As mentioned in the previous section, the dynamo model establishes that the magnetic field produced inside the star reaches the surface with the help of convective processes taken place in the CZ and gives rise to magnetically active regions on they reach the solar surface. Among them, we can find sunspots dominated by magnetic fields with an approximate intensity of $B\approx10^3\, G$ and coronal loops with field strengths of $B\approx10^2\, G$ at the photosphere, and $B\approx 10\, G$ in larger coronal heights \citep{Aschwanden2014}. Besides, rotating stars that expose a significant outer CZ can produce surface magnetic fields through a dynamo mechanism (see, e.g. \citet{Brandenburg2004}, \citet{Charbonneau2010} or \citet{Brun2017}). Whatever the origins of surface magnetic fields are, it's expected to couple to the wind mass-loss and, if this is strong enough, to produce MB \citet{UdDoula2002}, \citet{Ud-Doula2007}, \citet{Ud-Doula2008} \citet{Meynet2010}.\par

@@

Returning to section \ref{surf_mf} on the estimated mass loss ratio for a solar-type star, approximately $10^{-13}M_{\odot} \; yr^{-1}$, and using this value to estimate $\Dot{J}$, the result obtained can be considered relatively modest as to decisively influence the evolution of the star. On the other hand, if we consider that the convective movements that take place in the stellar interior together with the presence of differential rotation can give rise to the generation of magnetic fields that end up reaching the surface of the star \citep{Langer2012} and, the AML can be increased in several orders of magnitude as a consequence of the magnetic field that forces the ionized particles that compose the stellar wind to turn with the same angular speed and whose influence on the stellar wind extends a distance several times the radius of the star (see \citet{UdDoula2002}, \citet{Ud-Doula2007} and \citet{Ud-Doula2008} for more details).

This paper is structured as follows: Section 2 gives an overview of rotation and magnetic fields physical processes and how do they impact on the star evolution. We also summarized here the most relevant parameters configured in MESA for the model simulations. In Section 3, the model outputs are presented and discussed in detail. Finally, Section 4 finishes this paper with a discussion of open points and future works.\par

\section{Method}
\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA (based on \citet{Choi2016})}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Difussion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, $\alpha_{MLT}$ = 1.82\\
		Overshoot & time-dependent, diffusive, \\ & $f_{ov,core}=0.0160$,\\ 
		& $f_{ov,sh}=0.0174$\\
		Semiconvection & $\alpha_{sc}=0.1$\\
		Thermohaline & $\alpha_{th}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable between [3.5 - 5.5]\\
		Mass Loss & activated, $\Dot{M}_{max} = 10^{-3} \: M_{\sun} \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

In this section we review the core physic aspects simulated in MESA and the most relevant parameters adopted in the models simulations. In its simplest version, by activating rotation in the model, the evolutionary computational code MESA takes into account the following effects:
%\vspace{-1pc}
\begin{enumerate}
    \item AM transport from the radiative interior to the convective envelope in response to the rotational deceleration of the star surface layers\label{itm:1}.
    \item AM redistribution associated with changes in internal structure during the process of contraction to the MS\label{itm:2}.
    \item AML as a result of the torque applied to the convection zone by a magnetically coupled wind\label{itm:3}.
\end{enumerate}

@@

Keeping in mind this approach, our major objective is aimed to check the degree of validity of the resulting model after including the MB effect as a complementary mechanism of AML that can contribute to explain the evolution of Li in the Sun and other solar-type stars. A set of different simulation scenarios with magnetic field strengths ranging from between 3 and 5.5 G have been exercised. We'd like to stress again that the magnetic field remains constant during the PMS \& MS phases. We compute the evolution of $1M_{\sun}$ stellar models at solar metallicity with $\Omega / \Omega_{crit}$ varying between $0.0084$ and $0.0336$. More advanced models for more complex arrangements of the magnetic field as well as the evolution of its intensity throughout the life of the star are left for later work.\par

@@

It's also remarkable to observed, for some of the models depicted figure (\ref{fig:rot_vel_4g}), the lower $\Omega$ at the star surface (solid line) in comparison with the upper limit of the convective layer (dashed line). This fact is most probably a consequence of the expansion of the superficial layers after the mass loss produced by stellar winds.\par

>>>>

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS \& MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB (with magnetic field intensities ranging between 3.5 and 5.5 G) and the data obtained from the two set of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in star evolutionary models, impacts also the effective temperature ($T_{eff}$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the magnetic field intensity: the higher it is, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. Another important result of this study is that MB effect must be taken into consideration during PMS must if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

\section{Introduction} \label{sec_1}
Despite decades of theoretical efforts, a coherent explanation derived from models cannot be found for the Li abundance discrepancies detected in stars belonging to clusters of different ages and which are in the same evolutionary state, either in the PMS or in the MS. Additionally, these theoretical models are not able to explain the abundances detected in the late stages of the MS \citep{Tschape2001}.\par

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $T_{Li} \approx 2.5 x 10^6\; K$, and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. The Sun in particular and the solar-type stars in general are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $T_{Li}$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS) when the convection zone retreats and the temperature at BCZ is cooler than $T_{Li}$. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $T_{Li}$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances. Furthermore, they should also show a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this same period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the convective envelope, from its lower limit till the star surface. However, different abundances of Li have been observed for different star populations (\citet{Somers2014} and references therein).\par

From approximately the 1920s of the last century it has been accepted that rotation is capable of inducing the mixing of the elements present inside the stars. The studies of Eddington (1925) and Vogt (1925) had revealed the fact that the southern circulation must necessarily arise in those stars in rotation as a consequence of the von Zeipel paradox: a star cannot be simultaneously in thermal and hydrostatic equilibrium if its rotation speed depends exclusively on its radius. The solution to this depends on the existence of southern circulations. At the same time, this type of instability implies an impact on the convection mechanisms not considered by the standard models. Today it is known that in combination with this type of instability exists a catalogue of others that appear as a consequence of assuming models in rotation \citep{Maeder2003a}, particularly the differential rotation as a function of the radius of the star or the deviation of the spherical symmetry. Although initially these effects could be candidates to explain the observations, the fact is that the Sun is a slow rotator, meaning slow rotation that one for which the maximum centrifugal force is much smaller than the force of gravity. Furthermore, its deviation of the measured spherical symmetry is of the order of $10^{-5}$ and the time scale for the southern circulation is of the order of $10^{12}$ years, a time period well above the nuclear combustion time scale \citep{Pinsonneault1997}. In view of these findings, it seems evident that there are still open questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models.\par

The impact of rotation both on PMS and Li depletion for solar-type stars has already extensively debated in the past (\citet{Pinsonneault1997}, \citet{Jeffries2004}, \citet{Somers2014}) and revised more recently on the basis of the availability of more accurate measures (\citet{Bouvier2016}). However, these previous studies focused mainly on the hydrostatic effects and didn't consider the influence of a coupling between the star magnetic field and its possible spin-down effect. It's thought that AML has a direct influence in the mixing processes and that transport of the momentum could be produced by a number of relevant mechanisms: mass loss, magnetic fields and gravity waves (aka, g-modes). Concerning the two latter, gravity waves \citep{Charbonnel2005} and magnetic fields \citep{Eggenberger2009} have the property of transmiting angular momentum (AM) much more effectively than inducing mixing \citep{Denissenkov2007}. As a consequence of this increment in the AM transport efficiency, the amount of differential rotation between the radiative and the convective zones of the star is reduced (a solid-body rotation is fostered), as are the induced rotational instabilities. Some relevant efforts have already been undertaken aimed to model their link with AM distribution (\citet{Eggenberger2008}, \citet{Fuller2014}). As result, the efficiency of these processes is most easily tested by their influence on AM evolution. Of particular interest for this work is the impact caused by the magnetic field of the star in the AML, more specifically the effect produced on the solar wind particles that are forced to rotate in sync with it. As a consequence of this coupling a torque is induced which slows down the star, i.e. the magnetic braking effect. These magnetic fields would originate in the limits between the radiative and convective zones of the star, in the so-called tachoclines, due to the rotational difference between both zones \citep{Guerrero2016}. \par

Therefore, it's essential a proper consideration of the interactions between rotation and magnetic fields when it comes to study the AM distribution. We have adapted the one-dimensional stellar evolution code used in this work to model MB and included its effect on the equations which govern both AM distribution and stellar structure. This approach produces a more complete, though still reasonably simple, model for AM evolution. It's worth to mention that in this study we don't only focused in investigating the effects of rotation on the Li content during PMS and MS but also on the indirect role played by the MB on the Li destruction as a result of its influence on the rotational history of solar-type stars.\par

\subsection{Effects of rotation}
Rotation may affect the equations of stellar structure in four different ways according to \citet{Endal1976}:
%\renewcommand{\theenumi}{\arabic{enumi}.}
\begin{enumerate}
    \item The centrifugal force causes the local effective gravity ($g_{eff}$) associated with a point of the star not located on the axis of rotation to be reduced.
    \item In general, the centrifugal force will not be parallel to the gravitational force so that the equipotential layers will cease to be spherical.
    \item As the von Zeipel theorem states, the radiative flux in a uniformly rotating star is proportional to the $g_{eff}$, therefore the radiative flux is not constant on an equipotential surface.
    \item Rotation can suppress certain mechanisms that trigger convection, indirectly affecting mixing processes.
\end{enumerate}

Observational data support the fact that stars are born with different angular velocities even if they come from the same interstellar cloud. They also back up that they undergo a deceleration process throughout their evolution \citep{Skumanich1972} in a time period between 0.5 and 5 Gyr \citep{Somers2014a}. This range of different velocities implies that the mixing mechanisms induced by rotation vary their efficiency and effectiveness, and therefore have a direct influence on the abundances of the different chemical elements observed on the surface of stars. This fact, together with the differential degree of rotation to which the limits between the radiative and convective zones of the star are subjected, is specially interesting for our study since Li, as mentioned in the introduction, is destroyed at a low temperature. Therefore, variations in the intensity of the mixing processes can cause convective zones to extend to interior zones with temperatures higher than $T_{Li}$. The deceleration process to which a star is subjected throughout its evolution and the direct link that the rotation has with the mixing processes also offer a natural and consistent explanation for the fact that the destruction of Li is more accentuated or efficient in young stars than in old stars \citep{Sestito2005}. In a similar way, stellar rotation rates show a large dispersion at ZAMS \citep{Stauffer1984}, providing in this way a consistent argument for explaining the variance of Li abundances among stars which have a similar mass and were formed within the same cloud. Although this approach is not a novelty at all, it is no less true that in those studies based on it the effect caused by MB over the rotation on the PMS and MS is often overlooked. \par

@@

In the remainder of this paper, we describe a physically sound yet computationally simple model extension to Modules for Experiments in Stellar Astrophysics star evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds. In its simplest form, the implementation of the model requires the specification of both the surface magnetic field strength ($B$) and the initial rotation rate ($\Omega$) at star surface. The models include already rotation during the PMS as there evidences that advocate for a strong established relationship between Li destruction and rotation on that phase (\citeauthor{Bouvier2016} \citeyear{Bouvier2016}, \citeyear{Bouvier2018}), and the numerical simulations trace the rotational history and A(Li) of a $1\, M_{\sun}$ star for a variety of initial values for $B$ and $\Omega$.\par

We define here also some conventions and assumptions adopted in the paper. For the models presented in this paper we adopt solar-scaled abundances and assumes the \citet{Asplund2009} protosolar birth cloud metallicity ($Z = Z_{\odot, proto} = 0.0142$) not the current photospheric metallicity ($Z = Z_{\odot, photosphere} = 0.0134$) as the reference value. We also adopt the following nominal values to express stellar properties values in SI units $R_{\odot} = 6.957x10^{10}\, cm$ and $M_{\odot} = 1.988x10^{33}\, g$ which are consistent with IAU resolutions \citep{Mamajek2015}. For a detailed description of the physics adopted in this paper, refer to \citet{Choi2016}. That work, in particular its section 4.1, has been used as a starting point for ours regarding the parameterization of the MESA project which is calibrated to reproduce the measured element abundances on the solar surface.\par

@@

The mass of stars plays a determining role in the existence or not of an external convective envelope. In turn, the presence of this convective region conditions the evolution of magnetic fields. Massive stars do not have external convective zones although they can expose a convective core. In these cases, the dominant type of energy transport in the outer zones of the star is radiative. On the other hand, the less massive stars such as the Sun, have convective covers and a radiative core. In both types of stars strong magnetic fields can be generated in the tachocline but with notable differences. For the former, these magnetic fields can survive for a long time but not reach the surface of the star. For the latter, these magnetic fields, although they are able to reach the surface of the star, decay quickly in time, in a few hundred years or even only in a few decades \citep{Chabrier2006}. It's therefore necessary a mechanism for generating magnetic field, e.g. a dynamo type, continued in time to explain its existence. The way in which this dynamo process is able to generate magnetic fields depends on different parameters and to this day this mechanism is still not fully accepted and is subject of controversy, e.g. \citet{Charbonneau2010}.\par

\subsubsection{Internal Magnetic Fields}
Our Sun is an active star as far as the magnetic field is concerned. According to the theoretical models this magnetic field is generated by a dynamo type effect in the tachocline, the intermediate zone between the radiative core and the CZ \citep{Aschwanden2014}, and has an intensity of approximately $B\approx10^5\, G$. Additionally, these models predict that because of the differential rotation existing between the core and CZ, the lines of the magnetic field that at the moment of their generation have a poloidal orientation will be affected by the differential rotation and forced to adopt a toroidal typology. This change of typology in the magnetic field, from poloidal to toroidal, is known as the $\omega$-effect. On the other hand, we have that the opposite case can also occur, that is to say, starting from a toroidal magnetic field to obtain a poloidal configuration. This would be possible on the one hand, thanks to the convective movements of material that are produced in the CZ of the star and on the other, to assume that this material behaves as a perfect conductor. This second condition allows to assume that the magnetic field lines are "frozen" in the conductive fluid and therefore must move in solidarity with it. Under these circumstances the magnetic field lines will be pushed to higher areas of the CZ as the material moves vertically and, due to the combined effect of stellar rotation plus Coriolis forces, the magnetic lines will be twisted and give rise to a poloidal magnetic field. This is what is known as the $\alpha$-effect.\par

A number of theoretical models have been proposed based on $\alpha$ and $\omega$ effects as an explanation to the generation of magnetic fields, among which is the so-called $\alpha\omega$ or also known as dynamo effect which was proposed by \citet{Spruit2002}. By means of this model it is possible to explain the change in polarity of the magnetic field that occurs in the Sun every 11 years, completing a full cycle every 22 years. It's also worth to mention that a theoretical debate on this topic is still ongoing \citep{Denissenkov2007} and with regards to MESA, an implementation according to the work of Spruit-Tayler has been coded \citep{Paxton2013}.\par

\subsubsection{Surface Magnetic Fields} \label{surf_mf}
The intensity of the magnetic field measured on the Sun's surface is about $B\approx1\, G$, twice the value of the Earth's one. Additionally its topology is varied and complex. As mentioned in the previous section, the dynamo model establishes that the magnetic field produced inside the star reaches the surface with the help of convective processes taken place in the CZ and gives rise to magnetically active regions on they reach the solar surface. Among them, we can find sunspots dominated by magnetic fields with an approximate intensity of $B\approx10^3\, G$ and coronal loops with field strengths of $B\approx10^2\, G$ at the photosphere, and $B\approx 10\, G$ in larger coronal heights \citep{Aschwanden2014}. Besides, rotating stars that expose a significant outer CZ can produce surface magnetic fields through a dynamo mechanism (e.g. \citet{Brandenburg2004}, \citet{Charbonneau2010} or \citet{Brun2017}). Whatever the origins of surface magnetic fields are, it's expected to couple to the wind mass-loss and, if this is strong enough, to produce MB (e.g. \citet{UdDoula2002}, \citet{Ud-Doula2007}, \citet{Ud-Doula2008} \citet{Meynet2010}) .\par

@@

The AML loss caused by MB depends directly on the amount of mass lost by the star due to the stellar winds. The estimated mass loss ratio for a solar-type star, approximately $10^{-13}M_{\odot} \; yr^{-1}$, and the resulting AML derived from that value (refer to \ref{mod_mb} for details) can be considered relatively modest as to decisively influence the evolution of the star. Furthermore, the combined effect of convective movements and differential rotation lead to the generation of magnetic fields. Those magnetic fields end up reaching the surface of the star \citep{Langer2012}, so that the AML can be increased in several orders of magnitude. That increment is as a consequence of the magnetic field that forces the stellar wind ionized particles to turn with the same angular speed, winds which extends a distance several times the radius of the star (see \citet{UdDoula2002}, \citet{Ud-Doula2007} and \citet{Ud-Doula2008} for more details).

This paper is structured as follows: Section \ref{sec_2} gives an overview of rotation and magnetic fields physical processes and how do they impact on the star evolution. We also summarized here the most relevant parameters configured in MESA for the model simulations. In Section \ref{sec_3}, the model outputs are presented and discussed in detail. Finally, Section \ref{sec_4} finishes this paper with a discussion about open points and future works.\par

\section{Method} \label{sec_2}
\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA (based on \citet{Choi2016})}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Difussion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, $\alpha_{MLT}$ = 1.82\\
		Overshoot & time-dependent, diffusive, \\ & $f_{ov,core}=0.0160$,\\ 
		& $f_{ov,sh}=0.0174$\\
		Semiconvection & $\alpha_{sc}=0.1$\\
		Thermohaline & $\alpha_{th}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable between [3.5 - 5.5]\\
		Mass Loss & activated, $\Dot{M}_{max} = 10^{-3} \: M_{\sun} \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

In this section we review the core physic aspects simulated in MESA and the most relevant parameters adopted in the numerical simulations. In its simplest version, by activating rotation in the model, the evolutionary computational code MESA takes into account the following effects:
%\vspace{-1pc}
\begin{enumerate}
    \item AM transport from the radiative interior to the convective envelope in response to the rotational deceleration of the star surface layers\label{itm:1}.
    \item AM redistribution associated with changes in internal structure during the process of contraction to the MS\label{itm:2}.
    \item AML as a result of the torque applied to the convection zone by a magnetically coupled wind\label{itm:3}.
\end{enumerate}

@@

Keeping in mind this approach, our major objective is aimed to check the degree of validity of the resulting model after including the MB effect as a complementary mechanism of AML that can contribute to explain the evolution of Li in the Sun and other solar-type stars. A set of different simulation scenarios with magnetic field strengths ranging from between 3 and 5.5 G have been exercised. We'd like to stress that the magnetic field remains constant during the PMS \& MS phases. We compute the evolution of $1M_{\sun}$ stellar models at solar metallicity with $\Omega / \Omega_{crit}$ varying between $0.0084$ and $0.0336$. More advanced models for more complex arrangements of the magnetic field as well as the evolution of its intensity throughout the life of the star are left for later work.\par

@@

It's also remarkable to observed, for some of the models depicted in figure (\ref{fig:rot_vel_4g}), the lower $\Omega$ at the star surface (solid line) in comparison with the upper limit of the convective layer (dashed line). This fact is most probably a consequence of the expansion of the superficial layers after the mass loss produced by stellar winds.\par
